\bf Supervised Tensor Decomposition with Side Information on Multiple Modes

### 2020-09-04 tensor_regress/jasa-latexsource/tensor-jasa.tex
upload

\if1\blind
{
  \title{\bf Supervised Tensor Decomposition with Covariates on Multiple Modes}
  \author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
    Department of Statistics, University of Wisconsin-Madison}
  \maketitle
} \fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Exponential Family Tensor Regression with Covariates on Multiple Modes}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
We consider the problem of supervised tensor decomposition given covariates on multiple modes. Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling. We propose a new family of tensor response regression models that incorporate covariates, and establish the theoretical accuracy guarantees. Unlike earlier methods, our estimation allows high-dimensionality in both the tensor response and the covariate matrices on multiple modes. An efficient alternating updating algorithm is further developed. Our proposal handles a broad range of data types, including continuous, count, and binary observations. Through simulation and applications to two real datasets, we demonstrate the outperformance of our approach over the state-of-art.
\end{abstract}

\noindent%
{\it Keywords:}  Tensor-response models, Multiple covariates, Dimension reduction, Reduced-rank regression, Generalized linear models 
\vfill

@@

Nowadays, multi-dimensional arrays, a.k.a.\ tensors, are collected with additional covariates on multiple modes in many modern scientific and engineering studies. One example is neuroimaging~\citep{sun2017store,zhou2013tensor}. The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a). Another example is in the field of network analysis~\citep{baldin2018optimal,hoff2005bilinear}. A typical social network consists of nodes that represent people and edges that represent the friendships. Other covariates such as people’s demographic information and the closeness of a friendship are often available. In both examples, it is of keen scientific interest to identify the variation in the response (e.g., brain connectivity, social community patterns) that may be affected by additional covariates. These seemingly different problems can be formulated as a regression problem in the tensor sense.

@@

\begin{figure}[t]
\centering
\includegraphics[width=15cm]{loglike}
\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c)Poisson models. The dashed line represents the objective value at true parameter $\tL_{\tY}(\tB_{true})$. A micro-iteration refers to the update for one of the $K$ blocks. Four micro-iterations consist of a complete ``while'' iteration in Algorithm~\ref{alg:B}.}\label{fig:loglike}
\end{figure}

@@

\subsection{Nations data analysis}
Table~\ref{tab:s1} summarizes the $K$-means clustering of the 56 relations based on the $3^{\text{rd}}$ mode factor $\mM_3\in\mathbb{R}^{56\times 4}$ in the tensor regression model.

>>>>

\if1\blind
{
  \title{\bf Supervised Tensor Decomposition with Side Information on Multiple Modes}
  \author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
    Department of Statistics, University of Wisconsin-Madison}
  \maketitle
} \fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf General Tensor Regression with Covariates on Multiple Modes}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Higher-order tensors have recently received increasing attention across science and engineering. While most tensor decomposition methods are developed for a single tensor observation, scientific applications often generate side information, including node features and interactions thereof, together with the tensor data. 
Identifying the relationship between a high-dimensional tensor and multi-mode features is an important yet challenging problem in neuroimaging, network analysis, and spatial-temporal modeling. 
Here, we develop a regression approach that incorporates side information as structured covariates, and present several case studies to show the benefit of proposed approach.  
Unlike unsupervised tensor decomposition, our supervised tensor factors capture the most effective dimension reduction of the data tensor confined to the feature spaces. 
%supervised tensor decomposition method that incorporates side information on multiple modes. 
%We consider the problem of decomposing a higher-order tensor with exponential-family entries, given side information on multiple modes. 
%Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling. 
%We develop a regression approach that incorporates side information as structured covariates along multiple sides. 
%Unlike unsupervised tensor decomposition, the supervised approach finds the optimal low-rank factor matrices confined to the feature spaces. 
An efficient alternating updating algorithm is further developed. Our proposal handles a broad range of data types, including continuous, count, and binary observations. We apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data. Our approach identifies the key global connectivity pattern and pinpoints the local regions that are associated with covariates.

\end{abstract}

\noindent%
{\it Keywords:} Applications and case studies, Supervised dimension reduction, Exponential family observation, Reduced-rank tensor regression
\vfill

@@

Multi-dimensional arrays, known as tensors, are collected with additional covariates on multiple modes in many modern scientific and engineering studies. A popular example is the filed of neuroimaging~\citep{sun2017store,zhou2013tensor}. The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a). Another example is in the field of network analysis~\citep{baldin2018optimal,hoff2005bilinear}. A typical social network consists of nodes that represent people and edges that represent the friendships. Covariates such as people’s demographic information and the closeness of a friendship are often available. In both examples, it is of keen scientific interest to identify the variation in the array data (e.g., brain connectivity, social community patterns) that is affected by additional covariates. These seemingly different problems can be formulated as a regression problem in the tensor sense.

@@

\begin{figure}[t]
\centering
\includegraphics[width=15cm]{new_log.pdf}
\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models. The dashed line represents the objective value at true parameter $\tL_{\tY}(\tB_{true})$. A micro-iteration refers to the update for one of the $K$ blocks. Four micro-iterations consist of a complete ``while'' iteration in Algorithm~\ref{alg:B}.}\label{fig:loglike}
\end{figure}

@@
